Nail: A practical interface generator for binary formats

### 2014-04-01 paper/intro.tex
checkpoint

\subsection{Problems handling untrusted input}

Code handling untrusted user inputs is most prone to contain security
vulnerabilities. While memory corruption errors or fatal logic flaws
can occur anywhere in a program, bugs in the input handling code are
most likely to be exploitable, because an attacker has total control
over the inputs to that code.

Common wisdom thus says to take special care when preparing code that
deals with untrusted input.  However, with currently prevalent software
design methods, it is hard to isolate input-handling code from the rest
of the program, as user input is typically passed through the program and
processed in bits and pieces as a 'shotgun parser'~\cite{shotgun-parser}.

A look at the CVE database or Hacker journal such as
\textit{Phrack} shows that such design patterns and the security
problems they cause are rampant. Input processing libraries such as
libpng\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}
shows 24 remote exploits in the between 2007 and 2013.} or
Adobe's PDF and Flash viewers are notoriously plagued with
input bugs. Even simpler formats such as the \textit{zlib}
compression library have historically contained multiple security
flaws\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}.

Besides memory corruption vulnerabilities - against which various
mitigations such as static analysis, dynamic instrumentation and
various memory protection mechanisms have been developed - different
implementations of the same format can also differ in their understanding
of some edge cases. The security of many systems relies on different
components interpreting the same data similarly.

Such parser differential vulnerabilities were found in various
cryptographic systems, including  iOS\footnote{The XNU kernel and the
user code-signing verifier interpret executable metadata differently,
so the code signature sees different bytes at a virtual address than
the executable that runs.}~\cite{evaders6} and Android\footnote{Android
applications are distributed as .zip files. Signatures are verified
with a Java program, but the program is extracted with a C program.
The java program interprets all fields as signed, whereas the C
program treats them as unsigned, allowing one to replace files in a
signed archive.}~\cite{saurik-masterkey} code signing and even the
X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Such
vulnerabilities occur even if the language, runtime environment or static
analysis guarantee memory safety.

Finally, many protocol implementations are so complicated that even
without bugs, they inadvertently provide a full Turing-complete execution
environment. Examples include X86 page tables~\cite{bangert2013page},
ELF symbols and relocations~\cite{shapiro2013weird}. Even though
these protocols were never intended to be that expressive -
and consequentially hard to parse, implementation details of their
recognisers - which de facto became a part of the specification and were
relied upon by legitimate inputs - expose unwanted computational power
and have unintended side-effects.  In the offensive research community,
this has been generalised into treating a program as a \textit{weird
machine}~\cite{bratus2011exploit} that operates on an input, analogous
to a virtual machine operating on bytecode.

Proper input recognition has been shown to be an excellent way of
eliminating malicious inputs. In one case, a  PDF parser implemented in
Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of
file formats  and protocols in existence and might result in parser code
tied to one specific application.

\subsection{Parser generators}

>>>>

Code that handles untrusted inputs, such as processing network
data or parsing a file, is error-prone and is often exploited by
attackers.  This is in part because attackers have precise control
over the inputs to that code, and can craft inputs that trigger
subtle corner cases in input processing.  For example, the libpng
library has had 24 remotely exploitable vulnerabilities from 2007 to
2013,\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}}
and Adobe's PDF and Flash viewers have been notoriously
plagued by input processing vulnerabilities.  Even relatively
simple formats, such as those used by the zlib compression
library, have had input processing vulnerabilities in the
past.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}

### 2014-04-01 paper/relwk.tex
checkpoint

@@

>>>>

\paragraph{Parsing frameworks.}

@@

\paragraph{Language safety.}

Input processing vulnerabilities fall into two broad classes.  The first
class is memory safety bugs, such as buffer overflows, which allow an
adversary to corrupt the application's memory using specially crafted
inputs.  These mistakes arise in lower-level languages that do not
provide memory safety guarantees (such as C), and can be partially
mitigated by a wide range of techniques, such as static analysis,
dynamic instrumentation, and address space layout randomization, that
make it more difficult for an adversary to exploit these bugs.

The second class is logic errors, where application code misinterprets
input data.  This can lead to serious security consequences when two
systems disagree on the meaning of a network packet or a signed message,
as in iOS\footnote{The XNU kernel and the user code-signing verifier
interpret executable metadata differently, so the code signature
sees different bytes at a virtual address than the executable that
runs.}~\cite{evaders6} and Android\footnote{Android applications
are distributed as .zip files. Signatures are verified with a Java
program, but the program is extracted with a C program.  The java
program interprets all fields as signed, whereas the C program
treats them as unsigned, allowing one to replace files in a signed
archive.}~\cite{saurik-masterkey} code signing and even the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
highly application-specific, and are difficult to mitigate using existing
techniques, and these mistakes can occur even in high-level languages
that guarantee memory safety.

--- XXX integrate the following paragraphs somewhere into the logical flow ---

With currently prevalent software design methods, it is hard to isolate
input-handling code from the rest of the program, as user input is
typically passed through the program and processed in bits and pieces as a
``shotgun parser''~\cite{shotgun-parser}.

Finally, many protocol implementations are so complicated that even
without bugs, they inadvertently provide a full Turing-complete execution
environment. Examples include X86 page tables~\cite{bangert2013page},
ELF symbols and relocations~\cite{shapiro2013weird}. Even though
these protocols were never intended to be that expressive -
and consequentially hard to parse, implementation details of their
recognisers - which de facto became a part of the specification and were
relied upon by legitimate inputs - expose unwanted computational power
and have unintended side-effects.  In the offensive research community,
this has been generalised into treating a program as a \textit{weird
machine}~\cite{bratus2011exploit} that operates on an input, analogous
to a virtual machine operating on bytecode.

Proper input recognition has been shown to be an excellent way of
eliminating malicious inputs. In one case, a  PDF parser implemented in
Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of
file formats  and protocols in existence and might result in parser code
tied to one specific application.
